Preserving Reciprocal Consistency in Distributed Graph Databases

### 2020-04-10 paper/sec4-corruption.tex
camera-ready

\section{Corruption in the Absence of Concurrency Control}
\label{sec:db-corruption}

Consider the edge \texttt{(a)-[:\textcolor{green}{WROTE}]->(b)} spans servers $S_i$ and $S_j$.
When a transaction writes a distributed edge there are two important facts to consider:
\begin{enumerate}
\item Two write operations are performed, writing reciprocal entries in the adjacency lists of \emph{a} and \emph{b}.
\item Write order is unconstrained, a transaction is equally likely to write \emph{a} then \emph{b} as it is to write \emph{b} then \emph{a}.
\end{enumerate}

Without concurrency control, transactions $T_x$ and $T_y$ that write the distributed edge can interleave in the following ways:
\begin{enumerate}[(i)]
\item $T_x$ begins writing the distributed edge before $T_y$ but at opposite ends, transactions cross in the network and at $S_i$, $T_x \rightarrow T_y$ and at $S_y$, $T_y \rightarrow T_x$, Figure \ref{fig:a}.
\item $T_y$ begins writing the distributed edge before $T_x$ but at opposite ends, transactions cross in the network and at $S_i$, $T_x \rightarrow T_y$ and at $S_j$, $T_y \rightarrow T_x$, Figure \ref{fig:b}.
\item $T_x$ and $T_y$ perform their first writes at $S_i$, $T_x \rightarrow T_y$, but overlap in the network and arrive out-of-order at $S_j$, $T_y \rightarrow T_x$, Figure \ref{fig:c}.
\end{enumerate}

Each interleaving in Figure \ref{fig:conf} results in a \emph{half-corrupted} distributed edge - reciprocal consistency has been violated.
A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000})
To illustrate the process of half-corruption, consider $T_x$ deletes the \emph{wrote} edge and $T_y$ appends a property \emph{year}:
\input{cypher/delete-edge.tex}
\input{figures/hc}
Each interleaving results in the distributed edge in Figure \ref{hc-edge}, reciprocal consistency is clearly violated.
We say a graph with half-corrupted edges has suffered \emph{structural corruption}.

Assume we know the correct state of the distributed edge to be \texttt{(a)-[:\textcolor{green}{WROTE}]->(b)} with \texttt{\textcolor{red}{year} = 1937}, then there exists correct and incorrect entries for edge, visible in Figure \ref{hc-db-rep}.
When a transaction reads an edge it does not check both entries that constitute the edge are reciprocal consistent.
Therefore, a subsequent transactions can read the incorrect entry of a half-corrupted edge and write further edges, introducing \emph{semantic corruption} into the database.
Further semantic corruption spreads by the same mechanism.
A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupted state, rendering the database of little practical use.

For example, $T_z$, introduces semantic corruption if it reads \emph{b} and adds a new \emph{wrote} edge to an ``unknown author'' vertex.

\input{cypher/spread.tex}

If the database provides the ANSI isolation level \textbf{Read Uncommitted}, the database consistently orders writes from concurrent transactions, which would prevent all interleavings in Figure \ref{fig:conf} and the spread of semantic corruption

>>>>

\section{Reciprocally Inconsistent Distributed Edges}
\label{sec:db-corruption}

Suppose that the edge \texttt{(a)-[:\textcolor{green}{WROTE}]->(b)} is a distributed edge, with vertices \emph{a} and \emph{b} in servers $S_i$ and $S_j , j\neq i$,  respectively.
When a transaction writes this edge $ab$,
\begin{enumerate}
\item two writes are performed: reciprocal entries in the adjacency lists of both \emph{a} and \emph{b} are updated, and
\item write order is unconstrained: a transaction is equally likely to write \emph{a} then \emph{b} as it is to write \emph{b} then \emph{a}.
\end{enumerate}

Concurrent transactions $T_x$ and $T_y$ can interleave in the following three ways and each one is depicted in Fig \ref{fig:conf}:
\begin{enumerate}[(a)]
\item $T_x$ starts before $T_y$; it writes $a$ at $S_i$ first and then proceeds to $S_j$ across the network; $T_y$ operates the other way round, beginning with $S_j$ and proceeding to $S_i$ (see Fig \ref{fig:a}). The net effect is:  $T_x \rightarrow T_y$ at $S_i$ and at  $T_y \rightarrow T_x$ at $S_j$, where  $T \rightarrow T'$ at $S$ denotes that $T$ \emph{precedes} $T'$ at server $S$.
\item Same as the previous case, except that $T_y$ starts earlier than $T_x$ (see Fig \ref{fig:b}).
\item Same net effect as in previous two cases, except that both $T_x$ and $T_y$ start their first writes at $S_i$, $T_x \rightarrow T_y$, but $T_y$ overtakes $T_x$ in reaching $S_j$ where $T_y \rightarrow T_x$ (see Fig \ref{fig:c}).
\end{enumerate}

We could envisage three more corresponding cases (a') - (c') where the roles of $T_x$ and $T_y$ in cases (a) - (c) are simply interchanged; e.g., in case (a'), $T_y$ starts before $T_x$, writes $a$ at $S_i$ first and then proceeds to $S_j$ across the network; $T_x$ operates the other way round, beginning with $S_j$ and proceeding to $S_i$. Thus, there are only 6 ways concurrent $T_x$ and $T_y$ can interleave. The arguments we make based on cases (a) - (c) of Fig \ref{fig:conf} equally apply, by symmetry, to cases (a') - (c') and so, for brevity, we will not consider the latter.

At the end of each case in Fig \ref{fig:conf}, the last update on $a$ is by $T_y$ and that on $b$ is by $T_x$. Unless updates of $T_x$ and $T_y$ are commutative, $ab$ cannot be reciprocally consistent. As an example, suppose that $T_x$ deletes the \emph{wrote} edge while $T_y$ concurrently appends a property \emph{year}:
\input{cypher/delete-edge.tex}
\input{figures/hc}
The interleaving patterns depicted in Fig \ref{fig:conf} leave $ab$ reciprocally inconsistent, as shown in Fig \ref{hc}.

When $T_x$ and $T_y$ do not interleave,  either $T_y \rightarrow T_x$ or $T_x \rightarrow T_y$ holds at both servers $S_i$ and $S_j$; in that case, the last update on \emph{both} $a$ and $b$ would be by either $T_x$ or $T_y$, respectively. That is, when transactions update at both ends of $ab$  in \emph{some} arbitrarily chosen but identical order, $ab$ is left  reciprocally consistent after each transaction's update.

When interleaving updates leave $ab$ reciprocally inconsistent, $ab$ can be said to have become \emph{half-corrupted} because if $T_y \rightarrow T_x$ is the chosen order between $T_y$ and $T_x$, then the edge pointer in $b$ of $ab$ is in error; otherwise, the edge pointer in $a$ is erroneous. Thus, a reciprocally inconsistent edge $ab$ certainly has a corrupt half but the question of which half is corrupt has been left open.

Suppose that a future transaction $T_w$ \emph{first} reads the edge pointer of reciprocally in consistent $ab$, say, at vertex $a$. (Note that when $T_w$ reads any edge, it does not check for reciprocally consistency). At that moment, $T_w$ (implicitly) chooses the order $T_x \rightarrow T_y$ and thereby invalidates the other order $T_y \rightarrow T_x$ that prevails at vertex $b$. Thus, from that moment onward, the $b$ end of edge $ab$ becomes the corrupt end.

If no transaction ever reads the edge pointer at vertex $b$, then the order $T_x \rightarrow T_y$ effectively prevails and the half-corruption of $ab$ remains invisible to the rest of the database. However, if $T_z$ is to subsequently read the edge pointer at $b$ and write another edge based on what it read, i.e., The Hobbit has unknown author, then it is introducing updates not consistent with what $T_w$ read earlier; it thus introduces \emph{semantic corruption} into the database. Further writes based on reading semantically corrupt data also spread corruption.

A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupt state. As stated earlier, we had shown that 10\%  of a distributed graph database can become semantically corrupt well within the system lifetime itself (\cite{Ezhilchelvan2018}, \cite{Webber2019}).

Two more relevant remarks on past works:
a half-corrupted edge is due to a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}) in the context of distributed graph databases. If the database provides the ANSI isolation level \emph{Read Uncommitted}, it will identically order the writes of concurrent transactions and this would prevent all interleaving patterns shown in Fig \ref{fig:conf} and thus avert half-corruption altogether.

### 2020-04-10 paper/sec5-delta.tex
camera-ready

A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only once the acquiring transaction has committed or aborted. To prevent deadlock a policy such as \texttt{NO_WAIT} deadlock detection is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.

@@

These differences lead to several advantages in the context of graph databases. Firstly, a subset of edges are traversed and modified with a high frequency e.g. critical sections of motorway in a road network, leading to high contention. Secondly, graph transactions tend to be longer-lived than transactions in relational databases. Waiting for earlier writes on edges by long running transactions to commit, these frequently accessed edges significantly limit concurrency and reduce throughput. With these concerns in mind we developed the \tDelta protocol, which aimed at preventing edges becoming half-corrupted and hence quashing the seed of corruption whilst keeping performance at an acceptable level.\footnote{The \tDelta protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}

@@

The \tDelta Protocol has five rules:
\begin{enumerate}
\item A transaction's update on an end pointer of an edge is initially tentative which would become permanent only if that transaction is permitted to commit.
\item A tentative write is possible if the end pointer is either in a permanent state or the immediately preceding tentative update was done at least $\Delta$ time before (where time is measured as per local clock).
\item If a transaction successfully performs all its tentative updates, then it is permitted to commit; otherwise, it must abort.
\item A transaction commits, when all its tentative updates are made permanent, e.g., using an atomic commitment protocol.
\item Tentative updates of an aborting transactions are ignored. An ignored tentative update can make a new transaction abort for up to $\Delta$ time after it was created; it is harmless thereafter and can be garbage collected at any time.
\end{enumerate}

@@

Let us define $\delta$ as the bound estimate on the interval that may elapse between a transaction updating one end of a distributed edge at one server and updating the other end of the same edge at another server.

Let $\Delta$ be chosen such that $\Delta > \delta$. Now consider the interleaving in Figure \ref{fig:tx-ty} and let $t_x$ be the (global) time when $T_x$ starts at server $S_i$; similarly $t_y$ be the time $T_y$ starts at server $S_j$. Note $t_y < t_x$, i.e. $(t_x - t_y) > 0$.

Say, $T_y$ reaches $S_i$ at time $t_y + d$, where $d$ is the actual time elapsed between completing a tentative write at one end and starting at another end, let us assume that $d \leq \delta$. When $T_y$ arrives at $S_i$ it will find a tentative write done at time $t_x$. In this case, $t_y + d - t_x = d - (t_x-t_y) < d \leq \delta < \Delta$; so, $T_y$ will abort, preventing writes from interleaving and half-corrupting the edge. A similar arguments can be made for the scenario in with Figure \ref{fig:ty-tx}. For the interleaving in Figure \ref{fig:same-side}, if $t_y - t_x > \Delta$, then $T_y$ cannot overtake $T_x$ at $S_j$.

Assume that $d > \Delta$, i.e., the estimate $\delta$ does not hold at this moment. In Figure \ref{fig:tx-ty} interfering updates are avoided only if $t_y + d - t_x = d - (t_x-t_y) < d \leq \delta < \Delta$ otherwise, an edge becomes half-corrupted. This holds for the interleaving in \ref{fig:ty-tx}. Thus, in the extreme case $t_x = t_y$, reciprocal consistency is not guaranteed if $d > \Delta$ when the former exceeds its upper bound estimate $\delta$. For Figure \ref{fig:same-side}, if $d$ for $T_x$ is larger than $\Delta$, reciprocal consistency can be violated.

This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency. The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.

In summary, the \tDelta protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded edge can become half-corrupted and process of corruption described in Section \ref{sec:db-corruption} can still occur. Two questions naturally arise regards the performance of the protocol. For different values of $\Delta$:
\begin{itemize}
\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
\item How many transactions are aborted as a result of the protocol?
\end{itemize}

%The two paragraphs (The protocol leverages... edge is lost) can be commented out.

%Figures 5 and 6 can be removed.

% The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end. It is then assumed that the network delay between two servers can be predicted to be some $\Delta$. All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed. From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time. Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$. An aborted transaction, aborts any and every previous tentative write that it may have successfully completed. This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.

% \begin{figure}[H]
%   \centering
%   \begin{tikzpicture}[node distance=2cm]
%     % s1
%     \draw [thick] (1.3,3) -- (1.7,3);
%     \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
%     % s2
%     \draw [thick] (3.3,3) -- (3.7,3);
%     \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};

%     \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
%     \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);

%     \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};

%   \end{tikzpicture}
%   \caption{An example of the \tDelta protocol preserving reciprocal consistency.}
%   \label{delta-abort}
% \end{figure}

% Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption. Figure \ref{corruption-again} displays an example when the \tDelta protocol fails to maintain reciprocal consistency. The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.

% \begin{figure}[H]
%   \centering
%   \begin{tikzpicture}[node distance=2cm]
%     \draw [thick] (1.3,3) -- (1.7,3);
%     \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
%     \draw [thick] (3.3,3) -- (3.7,3);
%     \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
%         \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
%     \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};

%     \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
%     \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
%   \end{tikzpicture}
%   \caption{An example of the \tDelta protocol failing to preserve reciprocal consistency.}
%   \label{corruption-again}
% \end{figure}

>>>>

A straightforward solution to preventing dirty writes is for transactions to take long duration write locks \cite{Berenson1995}, releasing them only after the acquiring transaction has committed or aborted. To prevent deadlock, a policy such as \texttt{NO_WAIT} deadlock avoidance is used, which was shown to be the optimal policy in a distributed, partitioned database \cite{Harding2017}.

@@

These differences lead to several advantages in the context of graph databases which are characterized by the following two aspects.
First, a subset of edges are traversed and modified with a very high frequency e.g. critical sections of motorway in a road network, leading to high contention.
Secondly, graph transactions tend to be longer-lived than those in other databases.
So, having to wait for earlier writes to terminate while transactions are long lived, will severely limit the scope for concurrent processing and reduce throughput in a highly contentious environment.

With these concerns in mind we developed the \tDelta protocol, which aims at preventing edges from becoming half-corrupted and hence quashing the seed of corruption whilst keeping performance at an acceptable level.\footnote{The \tDelta protocol is a concurrency control mechanism only for distributed edge updates; it does not concern itself with updates on  vertices.}

@@

The \tDelta Protocol has five rules:
\begin{enumerate}
\item A transaction's write on an end pointer of an edge is initially tentative which would become permanent only if that transaction is permitted to commit.
\item A tentative write is possible only if the end pointer is either in a committed state or the immediately preceding tentative write was done at least $\Delta$ time before (where time is measured as per local clock).
\item If a transaction performs all its tentative writes, then it is permitted to commit; otherwise, it must abort.
\item A transaction commits, when all its tentative writes are made permanent, e.g., by using an atomic commit protocol.
\item Tentative writes of an aborting transactions are ignored. An ignored tentative write can make a new transaction abort unnecessarily for up to $\Delta$ time after it was created; it is harmless thereafter and can be garbage collected at any time.
\end{enumerate}

@@

Let us define $\delta$ as the bound \emph{estimate} on the interval that may elapse between a transaction completing its update at one end of a distributed edge at one server and starting its update at the other end of the same edge at another server.

Let $\Delta$ be chosen such that $\Delta > \delta$.
Now consider the interleaving in Fig \ref{fig:b} and let $t_x$ be the (global) time when $T_x$ starts at server $S_i$; similarly $t_y$ be the time $T_y$ starts at server $S_j$.
Since $T_x$ starts after $T_y$ in Fig \ref{fig:b}, $t_x > t_y$, i.e. $(t_x - t_y) > 0$.

Say, $T_y$ reaches $S_i$ at time $t_y + d_y$, where $d_y$ is the \emph{actual} time elapsed between $T_y$ completing a tentative write at one end and starting at another end, let us assume that $d_y \leq \delta$.
When $T_y$ arrives at $S_i$ it will find a tentative write already done at time $t_x$.
In this case, $t_y + d_y - t_x = d_y - (t_x-t_y) < d_y \leq \delta < \Delta$; so, $T_y$ will abort, preventing writes from interleaving and half-corrupting the edge. Similar arguments can be made for the scenario in Fig \ref{fig:a}, if $d_x \leq \delta$ where $d_x$ is the \emph{actual} time elapsed between $T_x$ completing a tentative write at one end and starting at another end.
Note that $t_y > t_x$ and $T_x$ will abort because it will find out, on reaching server $S_j$ for its next tentative write, that $t_x +d_x - t_y = d_x - (t_y-t_x) < d_x \leq \delta < \Delta$. Finally, for the interleaving in Fig \ref{fig:c}, if $t_y - t_x > \Delta$, then $T_y$ cannot overtake $T_x$ at $S_j$, if $d_x \leq \delta$ holds.

Assume that $d_x > \delta$ or $d_y > \delta$ is possible, i.e., the estimate $\delta$ can fail to hold.
In Fig \ref{fig:b}, interleaving  writes are avoided only if $t_y + d_y - t_x = d_y - (t_x-t_y) < d_y < \Delta$; otherwise, reciprocal inconsistency can occur.
Similarly, interleaving  writes of Fig \ref{fig:a} are avoided only if $t_x + d_x - t_y = d_x - (t_y-t_x) < d_x < \Delta$; otherwise, reciprocal inconsistency can occur.
Thus, in the extreme case $t_x = t_y$ and reciprocal consistency is not guaranteed if $\Delta \leq $ max $\{d_x , d_y \}$.

For Fig \ref{fig:c}, interleaving  writes are avoided only if $(t_x + d_x) - (t_y + d_y)  < \Delta$; given that $(t_y - t_x) < \Delta$, when $d_x$ exceeds $ \delta$, the interleaving writes of Fig \ref{fig:c} are avoided only if $d_x \leq 2 \Delta$; otherwise, reciprocal inconsistency can occur.

In summary, the \tDelta protocol eliminates interleaving of transactions during edge writes, so long as $\Delta$ remains larger than the interval $d$ that elapses between a transaction completing its write at one end of a distributed edge and starting at the other end.
Since the exact value of $d$ taken by a transaction cannot be known in advance, its bound $\delta$ is estimated with the best effort and $\Delta$ is chosen to be $\Delta > \delta$.

The larger the value of $\Delta$ used, the more likely is that $\Delta > d$ holds and half-corruption and thereby operational corruption are averted; also, on the downside, the more likely is that non-interleaving transactions will find their tentative writes within $\Delta$ time of each other and the later ones choose to abort \emph{unnecessarily}.
In the extreme case, choosing a very large $\Delta$ ($\Delta \approx \infty $) totally eliminates any risk of reciprocal inconsistency but does not allow any tentative write until the preceding one is made permanent. It is equivalent to enforcing \texttt{NO_WAIT} policy, wherein a requesting transaction that finds the requested record being locked, must abort.

Our performance evaluation will therefore  measure the following two metrics for various values of $\Delta$:
\begin{itemize}
\item Time taken for 10\% of a large database to be corrupt,
\item Number of transactions aborted per second.
\end{itemize}
If $d$ is exponentially distributed with mean $1/ \mu$, the probability of $d$ exceeding $\Delta$ is $ e^{(-\mu \Delta)}$. The values of $\Delta$ chosen will explore a range of probabilities of $\Delta$ being exceeded.

### 2020-04-10 paper/sec6-modeling.tex
camera-ready

\section{Modeling}
\label{sec:modeling}

To assess the protocols impact on time to operational corruption the model developed in \cite{Ezhilchelvan2018} was extended\footnote{An implementation and empirical evaluation in an existing systems was not performed as such an evaluation would have been impractical (need to compare database state at end of experiment with the linearizable truth), slow (real time) and expensive (requiring many hours of storage and compute time)}. A summary of the model is now provided, before discussing the extensions.

The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second. To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write. To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$. For each type, a fraction $f$ are distributed edges and the remainder are local edges. The network delay between servers was assumed to be exponentially distributed with rate $\rho$. At any moment in time an edge can be in one of four states:
\begin{enumerate}
\item Local and clean.
\item Distributed and clean.
\item Half-corrupted distributed edge arising from a dirty write.
\item Semantically corrupted.
\end{enumerate}
% The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.

% \begin{figure}[ht]
%   \centering
%     \begin{tikzpicture}[node distance=2cm]
%       \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
%       \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
%       \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
%       \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
%       \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
%       \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
%       \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
%       \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
%       \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
%     \end{tikzpicture}
%     \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
%     \label{state-transitions}
% \end{figure}

Probabilities are then derived for a given read operation returning a correct answer (states 1, 2 or the correct record in state 3) and all the reads by a given transaction returning correct answers. Then the probability of edge becoming half-corrupted $q_i$, by a given transaction arriving at time $t$ and operating on edge of type $i$ is derived. These probabilities are used to construct transition rates $a_{i,j}$ between states, which are used to simulate the process of corrupting the database and obtain estimates for the time to operational corruption, $U_{\gamma}$. At time $0$, all edges are clean (free from corruption), when a certain fraction, $\gamma$, of all edges become semantically corrupted, the database itself is said to be operationally corrupt. The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.

% (Appendix A)\footnote{To simplify the derivation of the new conflict probability it was assumed that messages sent from the same servers do not arrive out of order i.e. the interleaving in Figure 3(c) cannot occur.}.

The \tDelta protocol impacts the rate of corruption by reducing the probability a transaction corrupts a distributed edge, $q^{new}_i$. Of interest, therefore, is: how large or small is the value of $U_\gamma$ for a given value of $\gamma$ under the \tDelta protocol? Moreover, to quantify the number of aborts, a second simulation which focused specifically on the subset of frequently accessed distributed edges was performed. The answers to these questions depends on several parameters characterizing the following systemic aspects:
\begin{itemize}
\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
\item \emph{Workload}. Measured as transactions per second (TPS). Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
\item \emph{Distributed Write Delays and Choosing $\Delta$}. The smaller the delays the less likely the bound $\Delta$ is violated. Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
\end{itemize}

>>>>

\section{Performance Evaluation Strategies}
\label{sec:modeling}

To assess the time to operational corruption, we reused the model that we developed in \cite{Ezhilchelvan2018} and adapted it for the \tDelta protocol.
For completeness, the model is briefly explained first before we discuss the adaptation.
We note that an empirical evaluation of the time to operational corruption using a real system would be impractical due to the sheer length of time and the cost it would take to run the experiment.
(For certain values of $\Delta$, the model predicts anywhere between 1-75 years for operational corruption! Details are to follow.)

The model of \cite{Ezhilchelvan2018} assumes that transactions arrive in a Poisson stream with rate $\lambda$ transactions per second (TPS, for short).
Each transaction performs a random number ($K$) of read operations and then updates a single edge.

To model a scale-free graph, edges in the distributed database are divided into $T$ types, $i = 1, 2, \ldots, T$.
Type-1 edges represent the most popular edges to be accessed by transactions and type-$T$ edges are the least popular. Popular edges are smaller in number compared to the less popular ones.
$N$ denotes the total number of edges in the database.
For all edge types, a fraction $f$ are distributed.

At time $0$, all edges are assumed to be clean (free from corruption); thereafter, an edge can be in one of four states:
\begin{enumerate}
\item Local and clean.
\item Distributed and clean.
\item Distributed and half-corrupted.
\item Local or distributed and semantically corrupted.
\end{enumerate}

The delay $d$ is the time interval that elapses between a transaction completing its tentative write at one end of a distributed edge and starting at another end. It constitutes the window for concurrent transactions to interleave as portrayed in Fig \ref{fig:conf}. It is assumed to be exponentially distributed with mean $ 1/ \mu$ (i.e., at rate $\mu$).

The model parameters now enable us to compute the probability $q_i$ that a transaction updating a clean type-$i$ distributed edge, leaves it half-corrupted, in the absence of any concurrency control mechanism. (Our earlier work in \cite{Ezhilchelvan2018} assumes no concurrency control). These probabilities, $q_i, 1 \leq i \leq T$, are then used to compute transition rates $a_{j,k}$ between state $j$ to $k$, $1 \leq j,k \leq 4$. (Since local and distributed edges are fixed, $a_{1,2} = a_{2,1} = 0$.)

The rates $a_{j,k}$ are in turn used to simulate the spread of corruption within the database to estimate the first passage time $U_{\gamma}$  for a fraction ${\gamma}$ of $N$ edges to enter the state 4. (When $\gamma N$ edges become semantically corrupt, the database becomes operationally corrupt.)  The reader is directed to \cite{Ezhilchelvan2018} and \cite{Webber2019} for a granular discussion of the initial model.

\paragraph{\tDelta Protocol Adaptation}
Our concurrency protocol reduces the probabilities $q_i, 1 \leq i \leq T$ and we compute $q^{new}_i$ as shown in Appendix \ref{sec:deriv-confl-prob}.
The new probabilities $q^{new}_i$ are used in the model and simulations of \cite{Ezhilchelvan2018} to estimate $U_{\gamma}$.

\paragraph{Number of aborts per second.} To evaluate this metric for various values of $\Delta$, a second simulation that focuses specifically on the subset of most frequently accessed distributed edges was performed.

Note that both metrics that we set out to evaluate will be influenced by several parameters that characterize the database and other aspects:
\begin{itemize}
\item \emph{Database Size}. Size is expressed by the total number of edges $N$, and the fraction $f$ of distributed edges.
\item \emph{Workload}. Measured as transactions per second (TPS). Significant for measuring $U_{\gamma}$ are: the fraction of this load that writes after reads and the number of reads that precede a write.
\item \emph{Distributed Write Delays and Choosing $\Delta$}. The smaller the delays the less likely the bound $\Delta$ is violated. Conversely, smaller $\Delta$ is the more likely the bound $\Delta$ is violated.
\end{itemize}

### 2020-04-10 paper/sec7-evaluation.tex
camera-ready

The graph analyzed consisted of seven edge types, $N_1=10^4, N_2=10^5, N_3=10^6, N_4=10^7,  N_5=10^8, N_6=10^9, N_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
The number of read operations per query is geometrically distributed starting at $2$, with an average of $15$, before a write.
In all edge types, a fraction $0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.
The delay $d$ between a transaction completing a tentative write at one end and starting at another end is exponential distributed with a mean of $5ms$.
The database is initial clean and considered to be corrupted when $10$\% ($\gamma = 0.1$) of all edges are corrupted.
The time taken until operational corruption $U$, is measured in days.
$U$ considered for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload of $\lambda = (10000, ..., 100000)$.
The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
For each $\Delta$ the probability that $d$ exceeds $\Delta$ is $P(D > \Delta) = 4.5 \time 10^{-5}, 3.1 \times 10^{-7}, 2.1 \times 10^{-9}$ respectively.

The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Figure \ref{time-to-corruption-results}.
Under no isolation, $U$ ranges between 50-500 days.
For $50ms$ this increases to 1-75 years. For $\Delta = 75, 100ms$ the time to corruption vastly exceeds lifetime of any system making data corruption resulting from half-corrupted edges of little practical concern.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{./figures/delta}
  \caption{Time until operational corruption.}
  \label{time-to-corruption-results}
\end{figure}

The number of aborts per second for $\Delta = 50, 75, 100ms$ are given in Figure \ref{aborts-results} for the most popular edge type, $N=10^4$. The simulation was ran for 10 seconds for a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$. For $\Delta = 50$ the abort rate varies between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{./figures/aborts}
  \caption{Number of aborts per second.}
  \label{aborts-results}
\end{figure}

>>>>

The following parameter choices are inline in with industry experiences.
The graph analyzed consisted of seven edge types, $n_1=10^4, n_2=10^5, n_3=10^6, n_4=10^7,  n_5=10^8, n_6=10^9, n_7=10^{10}$, totaling 11 billion edges, with access probabilities $p_1 =0.5, p_2 =0.25, p_3=0.13, p_4=0.06, p_5=0.03, p_6=0.02$ and $p_7 =0.01$; a graph of this size would have approximately 1 billion vertices.
The number of read operations before a write per query is geometrically distributed starting at $2$, with an average of $15$.
In all edge types, $f=0.3$ are distributed, the remainder are local; in proportion with a good graph partitioning algorithm.

The delay $d$ between a transaction completing a tentative write at one end and starting at another end is exponential distributed with a mean of $5ms$.
The database is initially clean and considered to be operationally corrupted when $10$\% ($\gamma = 0.1$) of all edges are semantically corrupted.
The time taken until operational corruption, $U$, is measured in days.
We consider a range of transaction arrival rates, $\lambda = (1000, ..., 10000)$; a typical graph workload comprises of 90\% read-only transactions and 10\% read-write transactions \cite{Angles2020}, hence the chosen range reflects a total workload $(10000, ..., 100000)$.
The following $\Delta$ values were considered $\Delta = 50, 75, 100ms$.
For each $\Delta$ the probability that $d$ exceeds $\Delta$ is $P(d > \Delta) = 4.5 \time 10^{-5}, 3.1 \times 10^{-7}, 2.1 \times 10^{-9}$ respectively.

The results for measuring the impact of $\Delta$ on the time until operational corruption are given in Fig \ref{time-to-corruption-results} (where the $y$-axis in log scale).
With no concurrency control, $U$ ranges between 50-500 days.
For $\Delta = 50ms$,  $U$ increases to 1-75 years. For $\Delta = 75, 100ms$ the time to corruption is significantly large as shown Fig \ref{time-to-corruption-results}.

To evaluate the number $\alpha$ of aborts occurred per second, simulations were run for 10 seconds for each arrival rate, $\lambda = (1000, ..., 10000)$.
Fig \ref{aborts-results} reports the fraction $\frac{\alpha }{\lambda}$ for various values of $\lambda$. This fraction is also the probability that an incoming transaction is aborted due to the $\tDelta$ protocol.
For $\Delta = 50$ ms, the abort probability is between $1-5\%$, this increases to between $1-7\%$ and $1-9\%$ for $\Delta = 75$ and $\Delta = 100$ respectively.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\linewidth]{./figures/delta}
  \caption{Time until operational corruption.}
  \label{time-to-corruption-results}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\linewidth]{./figures/aborts}
  \caption{Fraction of aborts.}
  \label{aborts-results}
\end{figure}
